---
output: html_document
editor_options: 
  chunk_output_type: console
---
Load up all the required libraries and the data. 

```{r}
library(tidyverse)
library(ggrepel)
library(brms)
library(haven)
library(ggdag)
library(ggrepel)
library(patchwork)

theme_set(theme_bw() + 
            theme(panel.grid = element_blank()))

options(scipen = 999)

d <- read_dta("final_conflict.dta")
```

Standardize and clean some require variables. 


```{r}
d <- d %>% 
  mutate(
    best_est = ifelse(is.na(best_est), 0, best_est),
    incidents = ifelse(is.na(incidents), 0, incidents),
    deaths_per_10000 = (best_est/TotalPopn) * 1e4,
  ) %>% 
  mutate(
    e = rethinking::standardize(elevation_max),
    f = rethinking::standardize(norm_forest * 100),
    p = rethinking::standardize(pov_rate * 100),
    ci = rethinking::standardize(incidents), #conflict incidences
    cd = rethinking::standardize(best_est),
    cdt = rethinking:: standardize(deaths_per_10000)#conflict deaths. 
  )
```

The general DAG is as follows:

Elevation --> Conflict Incidences (ci)
Elevation --> Forest Cover (f)
Forest Cover(f) --> Conflict Incidences. 

Let's draw some scatterplots.

```{r}
# Relationship between Forest Cover and Conflict Incidences.
d %>% 
  ggplot(aes(x = f, y = ci)) +
  geom_point(shape = 1, size = 3, color = "firebrick")+
  labs(x = "Forest Cover (standardized)",
       y = "Conflict Incidences (standardized)")

# Relationship between Elevation and Conflict Incidences.
d %>% 
  ggplot(aes(x = e, y = ci)) +
  geom_point(shape = 1, size = 3, color = "firebrick4")+
  labs(x = "Maxium Elevation (standardized)",
       y = "Conflict Incidences (standardized)")

# Relationship between Forested Proportion and Elevation

d %>% 
  ggplot(aes(x = e, y = f)) +
  geom_point(shape = 1, size = 3, color = "firebrick4")+
  labs(x = "Maxium Elevation (standardized)",
       y = "Proportion of Forests (standardized)")+
    geom_text_repel(data = . %>% 
                      filter(distname %in%c("Achham","Rukum",
                                            "Rolpa", "Dang")),
                    aes(label = distname))


# Looks very much like an inverted U


# Any relationship between Conflict Incidences (not deaths) and the poverty rate.

d %>% 
  ggplot(aes(x = p, y = ci)) +
  geom_point(shape = 1, size = 3, color = "firebrick4")+
  labs(x = "Poverty Rate(standardized)",
       y = "Conflict Incidences (standardized)")



```


# Let's now see the relationship between forest cover and Conflict Incidences using brms

```{r}
m1_cf <- brm(
  data = d,
  family = gaussian,
  ci ~ 1 + f,
  prior = c(
    prior(normal(0, 0.2), class = Intercept),
    prior(normal(0, 0.5), class = b),
    prior(exponential(1), class = sigma)
  ),
  iter = 2000, warmup = 1000, cores = 4, chains = 4,
  sample_prior = T
)

m1_ce <- brm(
  data = d,
  family = gaussian,
  ci ~ 1 + e,
  prior = c(
    prior(normal(0, 0.2), class = Intercept),
    prior(normal(0, 0.5), class = b),
    prior(exponential(1), class = sigma)
  ),
  iter = 2000, warmup = 1000, cores = 4, chains = 4,
  sample_prior = T
)

```

Let's sample some priors:

```{r}
prior_cf <- prior_draws(m1_cf)

prior_cf %>% 
  slice_sample(n = 50) %>% 
  rownames_to_column("draw") %>% 
  expand_grid(f = c(-2, 2)) %>% 
  mutate(ci = Intercept + b * f) %>% 
  ggplot(aes(x = f, y = ci)) +
  geom_line(aes(group = draw))

# Plot the fitted values agains the data.

fitted(m1_cf) %>% 
  data.frame() %>% 
  bind_cols(d) %>% 
  
  ggplot(aes(x = f, y = ci))+
  geom_point(shape = 1, size = 3, color = "firebrick4")+
  geom_smooth(stat = "identity",
              aes(y = Estimate, ymin = Q2.5, ymax = Q97.5),
              color = "firebrick4",
              fill = "firebrick4",
              alpha = 1/4)+
  labs(x = "Forest Cover (std)",
       y = "Conflict Incidences(std)")

```

This positive relationship between forest cover and conflict incidence may be spurious because elevation of a district does impact both : highly elevated areas have low forest cover, but highly elevated areas are also prone to more conflict incidences due to its strategic nature.

For this we will need a multiple regression.


```{r}
m1_cfe <- brm(
  data = d,
  family = gaussian,
  ci ~ 1 + f + e,
  prior = c(
    prior(normal(0, 0.2), class = Intercept),
    prior(normal(0, 0.5), class = b),
    prior(exponential(1), class = sigma)
  ),
  iter = 2000, warmup = 1000, cores = 4, chains = 4,
  sample_prior = T
)

```
 
Let's make a full-luxury bayes framework for the entire causal model.


```{r}
c_model <- bf(ci ~ 1 + f + e)
e_model <- bf(f ~ 1 + e)


m2 <- brm(
  data = d,
  family = gaussian,
  c_model + e_model + set_rescor(FALSE),
  
  prior = c(
    prior(normal(0, 0.2), class = Intercept, resp = ci),
    prior(normal(0, 0.5), class = b, resp = ci),
    prior(exponential(1), class = sigma, resp = ci),
    
    prior(normal(0, 0.2), class = Intercept, resp = f),
    prior(normal(0, 0.5), class = b, resp = f),
    prior(exponential(1), class = sigma, resp = f)),
  iter = 2000, warmup = 1000, seed = 5, cores = 4, chains = 4,
  file = "fits/m2"
  
)
```

From the scatterplot and even thinking, it seems that elevation and forest cover and non-linearly related. High elevation predicts high forests upto a point and then forested area decreases. 

Let's make two models : One linking elevation and forest_cover in a linear fashion and the other one is non-linear. Let's plot both sisde by side. 

```{r}
d <- d %>% mutate(
  e_s2 = e^2
)
# Linear Relationship

m1_fe <- brm(
  data = d,
  family = gaussian,
  f ~ 1 + e,
  prior = c(
    prior(normal(0, 0.2), class = Intercept),
    prior(normal(0, 0.5), class = b),
    prior(exponential(1), class = sigma)
  ),
  iter = 2000, warmup = 1000, cores = 4, chains = 4,
  sample_prior = T,
  file = "fits/m1_fe"
)

m1_fe_nl <- brm(
  data = d,
  family = gaussian,
  f ~ 1 + e_s2,
  prior = c(
    prior(normal(0, 0.2), class = Intercept),
    prior(normal(0, 0.5), class = b),
    prior(exponential(1), class = sigma)
  ),
  iter = 2000, warmup = 1000, cores = 4, chains = 4,
  sample_prior = T,
  file = "fits/m1_fe_nl"
)


```
Let's look at both these graphs. Fit 2 shows that we probably need to account for the non-linearity in the relationship between forest cover and maximum elevation. 

```{r}
fit_1 <- fitted(m1_fe) %>% 
  data.frame() %>% 
  bind_cols(d) %>% 
  
  ggplot(aes(x = e, y = f))+
  geom_point(shape = 1, size = 3, color = "firebrick")+
  geom_smooth(stat = "identity",
              aes(y = Estimate, ymin = Q2.5, ymax = Q97.5),
              color = "firebrick",
              fill = "firebrick",
              alpha = 1/4) +
  labs(x = "Max Elevation (std)",
       y = "Forest Cover (std)")

fit_2 <- fitted(m1_fe_nl) %>% 
  data.frame() %>% 
  bind_cols(d) %>% 
  
  ggplot(aes(x = e, y = f))+
  geom_point(shape = 1, size = 3, color = "firebrick")+
  geom_smooth(stat = "identity",
              aes(y = Estimate, ymin = Q2.5, ymax = Q97.5),
              color = "firebrick",
              fill = "firebrick",
              alpha = 1/4) +
  labs(x = "Max Elevation (std)",
       y = "Forest Cover (std)")

fit_1 | fit_2

```


Non-linear full luxury bayes

```{r}
e_model_nl <- bf(f ~ 1 + e_s2)

m2_nl <-  brm(
  data = d,
  family = gaussian,
  c_model + e_model_nl + set_rescor(FALSE),
  
  prior = c(
    prior(normal(0, 0.2), class = Intercept, resp = ci),
    prior(normal(0, 0.5), class = b, resp = ci),
    prior(exponential(1), class = sigma, resp = ci),
    
    prior(normal(0, 0.2), class = Intercept, resp = f),
    prior(normal(0, 0.5), class = b, resp = f),
    prior(exponential(1), class = sigma, resp = f)),
  iter = 2000, warmup = 1000, seed = 5, cores = 4, chains = 4,
  file = "fits/m2_nl"
  
)

  
```


Let's make predictor residual plots.

Normal Model residuals.

```{r}
rf <- data.frame(residuals(m1_fe))
d$rf <- r$Estimate

m3_cf <- brm(
  data = d,
  family = gaussian,
  ci ~ 1 + rf,
  prior = c(
    prior(normal(0, 0.2), class = Intercept),
    prior(normal(0, 0.5), class = b),
    prior(exponential(1), class = sigma)
  ),
  iter = 2000, warmup = 1000, cores = 4, chains = 4,
  sample_prior = T
)

rf_nl <- data.frame(residuals(m1_fe_nl))
d$rf_nl <- rf_nl$Estimate

m3_cf_nl <- brm(
  data = d,
  family = gaussian,
  ci ~ 1 + rf_nl,
  prior = c(
    prior(normal(0, 0.2), class = Intercept),
    prior(normal(0, 0.5), class = b),
    prior(exponential(1), class = sigma)
  ),
  iter = 2000, warmup = 1000, cores = 4, chains = 4,
  sample_prior = T
)

```


Plotting Predictor residual plots.

```{r}
f1 <- fitted(m3_cf) %>% 
  data.frame() %>% 
  bind_cols(d) %>% 
  
  ggplot(aes(x = rf, y = ci))+
  geom_point(shape = 1, size = 3, color = "red")+
  geom_smooth(stat = "identity",
    aes(y = Estimate, ymin = Q2.5, ymax = Q97.5),
    color = "red", fill = "firebrick", alpha = 1/4)+
  labs(x = "Forest Cover Residual (linear)")

f2 <- fitted(m3_cf_nl) %>% 
  data.frame() %>% 
  bind_cols(d) %>% 
  
  ggplot(aes(x = rf_nl, y = ci))+
  geom_point(shape = 1, size = 3, color = "red")+
  geom_smooth(aes(y = Estimate, ymin = Q2.5, ymax = Q97.5),
              stat = "identity",
              color = "red",
              fill = "firebrick",
              alpha = 1/4)


f1 | f2
```

